GMSE: an R package for generalised management strategy evaluation

### 2018-04-17 notebook/ms/SI3.Rmd
Edit parameters and include warning

@Johnson2018 recently estimated key demographic parameters for the Taiga Bean Geese for the Central Management Unit that includes geese that breed in "Northern most Sweden, Northern Norway, Northern and Central Finland, and adjacent North-wester parts of Russia, wintering mostly in Southern Sweden and South-east Denmark" [@AEWA2016]. They estimated goose survival under ideal conditions to be ca 0.878; this can be interpreted in our model by setting mortality to `remove_pr = 1 - 0.878`. Similarly, @Johnson2018 estimated intrinsic population growth rate and carrying capacity to be 0.15 and 93870, respectively, so we set `lambda = 0.15` and `res_death_K = 93870`. The global abundance of Taiga Bean Geese in 2009 was ca 63000 [@Fox2010], and ca 35000 in the Central Management Unit [@AEWA2016], which we can take as a starting abundance for our simulations (`RESOURCE_ini = 35000`). And the International Single Species Action Plan has a target population size of ca 70000 in the Central Management Unit [@AEWA2016], which we can use as a management target (`manage_target = 70000`).

@@

The code below calls `gmse` using the empirically derived parameters for Taiga Bean Geese described above. We also set `manager_budget = 100000` and `user_budget = 100000`. Further, we consider the case in of a region in which farmland makes up 60% of all land, with 40% of land being 'public' (`public_land = 0.4`; which might be intpreted as any land in which stakeholders are not, or cannot be, invested in goose presence), and divide the farmland amongst 80 individual farmers (`stakeholders = 80`; `land_ownership = TRUE`). Landscape size is set to default 100 by 100 cells, so each farmer effectively owns 75 cells, which might be interpreted as hectares of land (for instructions on how to more precisely control landscape ownership, see the advanced GMSE options using `gmse_apply`). Because we need both density-dependent (`res_death_K = 93870`) and density-independent (`remove_pr = 0.122`) sources of mortality, we set `res_death_type = 3`. We assume that a single goose decreases agricultural production on a cell by 1% per time step (`res_consume = 0.01`). We further assume that the population is well-monitored, with observers counting goose numbers on each cell of the landscape in every timestep (`observe_type = 3`) with the ability to observe two landscape cells in every direction (`agent_view = 2`). All other parameter values are set to GMSE defaults.

@@

Below, we first allow only the culling of geese by users and plot the dynamics of the social-ecological from a single simulation. Next, we run the same simulation but allow users to scare geese from their landscape cells; we then use the model to make inferences regarding how scaring as a management option might affect goose population dynamics, agricultural production, and conservation conflict in the system.

>>>>

@Johnson2018 recently estimated key demographic parameters for the Taiga Bean Geese for the Central Management Unit that includes geese that breed in "Northern most Sweden, Northern Norway, Northern and Central Finland, and adjacent North-wester parts of Russia, wintering mostly in Southern Sweden and South-east Denmark" [@AEWA2016]. They estimated goose survival under ideal conditions to be ca 0.878; this can be interpreted in our model by setting mortality to `remove_pr = 1 - 0.878`. Similarly, @Johnson2018 estimated intrinsic population growth rate and carrying capacity to be 0.15 and 93870, respectively, so we set `lambda = 0.183` (for convenience of demonstration, the upper 90% confidence interval of their estimate) and `res_death_K = 93870`. The global abundance of Taiga Bean Geese in 2009 was ca 63000 [@Fox2010], and ca 35000 in the Central Management Unit [@AEWA2016], which we can take as a starting abundance for our simulations (`RESOURCE_ini = 35000`). And the International Single Species Action Plan has a target population size of ca 70000 in the Central Management Unit [@AEWA2016], which we can use as a management target (`manage_target = 70000`).

@@

The code below calls `gmse` using the empirically derived parameters for Taiga Bean Geese described above. We also set `manager_budget = 100000` and `user_budget = 100000`. Further, we consider the case in of a region in which farmland makes up 60% of all land, with 40% of land being 'public' (`public_land = 0.4`; which might be intpreted as any land in which stakeholders are not, or cannot be, invested in goose presence), and divide the farmland amongst 80 individual farmers (`stakeholders = 80`; `land_ownership = TRUE`). Landscape size is set to default 100 by 100 cells, so each farmer effectively owns 75 cells, which might be interpreted as hectares of land (for instructions on how to more precisely control landscape ownership, see the advanced GMSE options using `gmse_apply`). Because we need both density-dependent (`res_death_K = 93870`) and density-independent (`remove_pr = 0.122`) sources of mortality, we set `res_death_type = 3`. We assume that a single goose decreases agricultural production on a cell by 2% per time step (`res_consume = 0.02`). We further assume that the population is well-monitored, with observers counting goose numbers on each cell of the landscape in every timestep (`observe_type = 3`) with the ability to observe two landscape cells in every direction (`agent_view = 2`). All other parameter values are set to GMSE defaults.

@@

Below, we first allow only the culling of geese by users and plot the dynamics of the social-ecological system from a single simulation. Next, we run the same simulation but allow users to scare geese from their landscape cells; we then use the model to make inferences regarding how scaring as a management option might affect goose population dynamics, agricultural production, and conservation conflict in the system. All simulations are run using the `gmse` function, but novice R users may prefer run all of the simulations below using the browser-based GMSE GUI by calling `gmse_gui()` from the R command line. Alternatively, experienced R users may prefer simulate by looping time steps through `gmse_apply`, which allows more flexibility for incorporating custom submodels and dynamically adjusting parameter values. We emphasise that the simulations below are intended only to demonstrate one use of GMSE on a species of conservation interest, not to make recommendations for management of Taiga Bean Geese.
